### 2013-08-30 thesis/src/chapters/chapter-2.tex
Added vertical scaling stub

@@

\subsection{Horizontal scaling}
As outlined in previous section, horizontal scaling is all about adding supplementary nodes to a system. As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion. As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later. While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack, OpenNebula, OpenStack, CloudStack or Eucalyptus to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done. Table 1 describes in details different capabilities of commonly used virtualization technologies.

Provisioning new server is only a first step in scaling an application, it is required to configure some load balancing mechanism to make user of this additional node.

@@

>>>>

\chapterintro{This chapter is devoted to the concept of scaling users' application from the perspective of a cloud platform provider. To achieve that, it presents attainments of research groups working in that area as well as it considers mechanisms used in products currently available on the market.}

@@

\subsection{Horizontal scaling}
As outlined in previous section, horizontal scaling is all about adding supplementary nodes to a system. As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion. As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later. While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done. Table \ref{tab:hypervisors-provisioning} describes in details different capabilities of commonly used virtualization technologies.

\begin{table}[!htbp]
\begin{tabularx}{\textwidth}{l | X | X}
 &  & \\
\hline 
KVM &  & \\
\hline
Xen &  & \\
\hline
VMware &  & \\
\hline
OpenVZ &  & \\
\hline
LXC &  &  \\
\end{tabularx}
\caption{Comparison of hypervisors provisioning capabilities}
\label{tab:hypervisors-provisioning}
\end{table}

Provisioning new server is only a first step in scaling an application, it is required to configure load balancing mechanism to make use of additional node. Two important aspects that has to be consider are: load-balancing algorithms and scalability.

\subsubsection*{Load-balancing algorithms}
Generally, there are two types of load-balancers: hardware and software based. Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility. Among the most common algorithms we can distinguish:
\begin{itemize}
 \item \textit{round-robin scheduling} -
 \item \textit{first in first out} -
 \item \textit{shortest remaining time} -
\end{itemize}

Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}. Table \ref{tab:load-balancers} summaries available load-balancing mechanism including their utilization in IaaS products.

\begin{table}[!htbp]
\begin{tabularx}{\textwidth}{l | X | X}
 &  & \\
\hline 
HAProxy &  & \\
\hline
Apache modproxy &  & \\
\hline
... &  & \\
\end{tabularx}
\caption{Comparision of load balancers}
\label{tab:load-balancers}
\end{table}

\subsubsection*{Load-balancing scalability}
Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer. In other words, load-balancer becomes a new SPOF. Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered. This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level \textbf{TODO check :)}

@@

Essentially, vertical scaling is concentrated upon increasing capacity of single node. Again, when considering technical advancements that comes with cloud computing and virtualization, we can differ two categories of scaling: virtual machine resizing and virtual machines replacement. This distinction is dictated by limitation hypervisors - not all of them are able to resize virtual machine without shutting it down, in that case provisioning new one is a necessity.

\subsubsection*{Virtual machine resizing}

\begin{table}[!htbp]
\begin{tabularx}{\textwidth}{l | X | X}
 &  & \\
\hline 
KVM &  & \\
\hline
Xen &  & \\
\hline
VMware &  & \\
\hline
OpenVZ &  & \\
\hline
LXC &  &  \\
\end{tabularx}
\caption{Comparision of hypervisors resizing capabilities}
\label{tab:hypervisors-resizing}
\end{table}

\subsubsection*{Virtual machine replacement}
